---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r echo = FALSE, results = "hide", message = FALSE}
library(badger)
```

# rethnicity

<!-- badges: start -->
```{r, echo = FALSE, results = "asis"}
cat(
  badge_cran_release("rethnicity", "green"),
  badge_cran_download("rethnicity", "grand-total", "green"),
  badge_cran_checks("rethnicity"),
  badge_last_commit("fangzhou-xie/rethnicity")
)
```
[![Build Status](https://app.travis-ci.com/fangzhou-xie/rethnicity.svg?branch=main)](https://travis-ci.com/github/fangzhou-xie/rethnicity)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

The goal of rethnicity is to provide a method to predict ethnicity from names of people.

## Installation

<!-- You can install the released version of rethnicity from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("rethnicity") -->
<!-- ``` -->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("fangzhou-xie/rethnicity")
```

## Getting to use the package in 5 minutes

```{r}
library(rethnicity)
```

### I want to predict ethnicity from last names!

Sure. I have trained a model to predict and classify race based on last names.
Simply use it as:

```{r}
predict_ethnicity(lastnames = "Freeman", method = "lastname")
```

### What if I have both first names and last names?

Of course. There is a separate model just to do that.
By having both first names and last names, we can achieve higher accuracy than only having last names.
The syntax is similar to what we have seen from above.

```{r}
predict_ethnicity(firstnames = "Morgan", lastnames = "Freeman", method = "fullname")
```

### What if I have multiple names?

Cool. I got you covered. Just use vectors as input.

```{r}
firstnames <- c("Morgan", "Will")
lastnames <- c("Freeman", "Smith")
predict_ethnicity(lastnames = lastnames, method = "lastname")
predict_ethnicity(firstnames = firstnames, lastnames = lastnames, method = "fullname")
```

Just remember to have the same length for the `firstnames` and `lastnames` vectors and the first name and last name
for the same person should have same index in each of the vectors.

### Wait. I want to predict a million names!

Alright. The package also supports extremely fast execution by multi-threading via the wonderful `RcppThread` package.
To use this, just pass a number to the `threads` argument and the number need to be greater than 0.

```{r}
lastnames <- rep("Freeman", 1000000)
# measure the elapsed time
start_time <- Sys.time()
p <- predict_ethnicity(lastnames = lastnames, method = "lastname", threads = 20)
end_time <- Sys.time()
end_time - start_time
```

For most use cases that I can imagine, the default setting (`threads = 0`) should be fast enough since we are leveraging C++
routines for the processing. If you have very large dataset, or if you have a powerful machine, or if you just
want to run the code faster, you can set the `threads` argument to be bigger than 0 and you should obeserve
performance boost.

Note that I have 24-thread desktop and hence choosing 20 threads to work on this. You may need to wisely choose
the appropriate number of threads for the job. In general, the more threads you have, the faster it should run.
But the relationship is not linear, as there will be more overhead when increasing the number of threads.

<!-- ```{r} -->
<!-- lastnames <- rep("Freeman", 1000000) -->
<!-- # measure the elapsed time -->
<!-- start_time <- Sys.time() -->
<!-- p <- predict_ethnicity(lastnames = lastnames, method = "lastname", threads = 10) -->
<!-- end_time <- Sys.time() -->
<!-- end_time - start_time -->
<!-- ``` -->

Processing a million names only spent around 10 seconds for us. I would call this pretty fast.
