---
title: "Introduction to `rethnicity` package"
output: rmarkdown::html_vignette
author: Fangzhou Xie
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

In this package, I aim to provide a function that could predict ethnicity (race) from names.

```{r setup}
library(rethnicity)
```

# WARNING!

I created this package hoping to help applied researchers on their studies regarding ethnic bias and discrimination,
and potentially eliminate the racial and ethnic disparities. By using this package, you agree to the following:

1. You **will NOT** use this package for purposes other than academic research.
2. You **will NOT** disclose the predicted ethnic group to the public, given the names data you might have.
3. You **will NOT** discriminate anyone on the basis of race and color, by using the methods provided by this package.
4. You **understand** that the method cannot make predictions 100% correct, and you should be cautious about the results.
<!-- 4. You **agree** to advocate racial equality. -->

Again, you should use the package responsibly.

# Getting to use the package in 5 minutes

## I want to predict ethnicity from last names!

Sure. I have trained a model to predict and classify race based on last names.
Simply use it as:

```{r}
predict_ethnicity(lastnames = "Jackson", method = "lastname")
```

## What if I have both first names and last names?

Of course. There is a separate model just to do that.
By having both first names and last names, we can achieve higher accuracy than only having last names.
The syntax is similar to what we have seen from above.

```{r}
predict_ethnicity(firstnames = "Samuel", lastnames = "Jackson", method = "fullname")
```

## What if I have multiple names?

Cool. I got you covered. Just use vectors as input.

```{r}
firstnames <- c("Samuel", "Will")
lastnames <- c("Jackson", "Smith")
predict_ethnicity(lastnames = lastnames, method = "lastname")
predict_ethnicity(firstnames = firstnames, lastnames = lastnames, method = "fullname")
```

Just remember to have the same length for the `firstnames` and `lastnames` vectors and the first name and last name
for the same person should have same index in each of the vectors.

## Wait. I want to predict a million names!

Alright. The package also supports extremely fast execution by multi-threading via the wonderful `RcppThread` package.
To use this, just pass a number to the `threads` argument and the number need to be greater than 0.

```{r}
firstnames <- rep("Samuel", 10000)
lastnames <- rep("Jackson", 10000)
# measure the elapsed time
start_time <- Sys.time()
p <- predict_ethnicity(firstnames = firstnames, lastnames = lastnames, threads = parallel::detectCores()-2)
end_time <- Sys.time()
end_time - start_time
```

Processing ten thousands names only spent around 0.64 seconds for us. I would call this pretty fast.

For most use cases that I can imagine, the default setting (`threads = 0`) should be fast enough since we are leveraging C++
routines for the processing. If you have very large dataset, or if you have a powerful machine, or if you just
want to run the code faster, you can set the `threads` argument to be bigger than 0 and you should observe
performance boost.

You may need to wisely choose the appropriate number of threads for the job. 
In general, the more threads you have, the faster it should run.
But the relationship is not linear, as there will be more overhead when increasing the number of threads.
In the example, I was choosing the number of threads by the maximum number minus 2 (24 - 2 = 22).

<!-- ```{r} -->
<!-- lastnames <- rep("Freeman", 1000000) -->
<!-- # measure the elapsed time -->
<!-- start_time <- Sys.time() -->
<!-- p <- predict_ethnicity(lastnames = lastnames, method = "lastname", threads = 10) -->
<!-- end_time <- Sys.time() -->
<!-- end_time - start_time -->
<!-- ``` -->


# FAQ

## How did you train the models?

I first trained the models in Keras with Python, using the Florida Voter Registration dataset.
After training a big model for the prediction, I also trained a smaller model than will mimic
the prediction of large model (this is called "distillation of knowledge").
By doing so, we could significantly reduce the size of the model while keeping the accuracy.
This is very important if we want the package to be lightweight and fast in processing data.

## How did you create the package with Keras models?

After the training and testing process, I save the distilled model and export it into C++ by the `frugally-deep` project.
This will allow us to get rid of the dependency on Keras and Python and we can directly making inferences
from C++. From here, it is very obvious that we can wrap the inference procedure by `Rcpp` and call it from R.

Note that one could potentially use the `keras` package in R to load Keras models trained in Python.
But I believe this would have defeated the purpose of having a R package, as the `keras` package
still depends on Python and the installation of Keras. You can argue that we are actually still using Python:)

## What is the difference between `fullname` model vs. `lastname` model?

<!-- TODO: potentially cite my article in later CRAN submission -->
I have trained two different model for predicting ethnicity from names, one only leverages last names, while
another incorporate both first names and last names. In some applications, researchers may 
only have access to last names, then they should consider using the `lastname` method.
In other cases, we could also have first names available, then we could incorporate this information
and use the `fullname` method. This will yield better accuracy for the prediction.

## What about the performance of the package?

The processing speed is super fast, as the heavy-lifting has been delegated to the underlying C++.
What is more, to make it even faster, I used `RcppThread` to achieve multi-threading.
This would be extremely useful if you have a very large dataset at hand.
As shown in the example above, we have achieved to process a million names within 10 seconds.
In other words, we could predict the race from a name by 10 μs on average.
